### 2020-05-16 writeup/proof_transfer_ex.tex

\subsection{Proof of Proposition \ref{prop_dist_transition}}\label{app_proof_31}

@@

Based on Lemma \ref{lem_hat_v}, we can prove Proposition \ref{prop_dist_transition}.
\begin{proof}[Proof of Proposition \ref{prop_dist_transition}]
	Since $\Sigma_1 = \Sigma_2$, we know that $M = \Sigma_1^{1/2}\Sigma_2^{-1/2} = \hat{v} \id$.
	Using Lemma \ref{lem_minv} and \ref{lem_cov_shift}, we can track the reduction of variance from $\hat{\beta}_t^{\MTL}$ to $\hat{\beta}_t^{\STL}$ as whp
	\begin{align*}
		\Delta_{\vari} &\define \sigma^2 \bigbrace{\frac{1}{c_2 - 1} - \frac{1}{c_1 + c_2}\cdot\frac{1}{a_1\hat{v}^2 + a_2} + \bigo{p^{-1/2+\varepsilon}}} \\
		&= \sigma^2 \bigbrace{\frac{c_1}{(c_2-1)(c_1 + c_2 -1)} + \bigo{p^{-1/2 + \varepsilon}}},
	\end{align*}
	where we use equation \eqref{eq_a12} and Lemma \ref{lem_hat_v}.
	Next we consider the model shift bias
	\begin{align*}
		\Delta_{\beta} &\define \hat{v}^2 \bignorm{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_s - \hat{v}\beta_t)}^2 \\
		&= d^2 \cdot \bignormFro{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1}^2 + \bigo{p^{-1/2 + \varepsilon}}
	\end{align*}
	Using Lemma \ref{prop_model_shift_tight}, we get an upper and lower bound on $\Delta_{\beta}$ as
	\[ \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 \le \Delta_{\beta} / \bigbrace{p\cdot d^2 \cdot \frac{c_1^2}{(c_1+c_2)^2} \cdot \frac{1 + a_3 + a_4}{(a_1 + a_2)^2} + \bigo{\frac 1 p} } \le \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4. \]
	By solving equations \eqref{eq_a12}, \eqref{eq_a3} and \eqref{eq_a4}, we get
	\begin{align*}
		& a_1 = \frac{c_1(c_1 + c_2 - 1)}{(c_1 + c_2)^2} + O(p^{-1/2 + \varepsilon}),
		a_2 = \frac{c_2(c_1 + c_2 - 1)}{(c_1 + c_2)^2} + O(p^{-1/2 + \varepsilon}), \\
		& a_3 = \frac{c_2}{(c_1 + c_2)(c_1 + c_2 - 1)} + O(p^{-1/2 + \varepsilon}),
		a_4 = \frac{c_1}{(c_1 + c_2)(c_1 + c_2 - 1)} + O(p^{-1/2 + \varepsilon}).
	\end{align*}
	Hence we obtain that
	\begin{align*}
		\frac{1 + a_3 + a_4}{(a_1 + a_2)^2}
		= \frac{(c_1 + c_2)^3}{(c_1 + c_2 - 1)^3} + \bigo{p^{-1/2+\varepsilon}}.
	\end{align*}
	To sum up, we have shown that when
	\[ \frac{p\cdot d^2}{\sigma^2} < \frac{(c_1 + c_2 - 1)^2}{c_1(c_1 + c_2) (c_2 - 1)} \cdot \bigbrace{1 + \sqrt{\frac 1 {c_1}}}^{-4}, \]
	we have that $\Delta_{\vari} > \Delta_{\beta}$, which implies that $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$.
	When
	\[ \frac{p\cdot d^2}{\sigma^2} > \frac{(c_1 + c_2 - 1)^2}{c_1 (c_1 + c_2) (c_2 - 1)} \cdot \bigbrace{1 - \sqrt{\frac 1 {c_1}}}^{-4}, \]
	we have that $\Delta_{\vari} < \Delta_{\beta}$, which implies that $\te(\hat{\beta}_t^{\MTL})$ is higher than $\te(\hat{\beta}_t^{\STL})$.
	The result follows by taking $c_1 = c_2 = c$ for the above equation.
\end{proof}

@@

>>>>

\subsection{Proof of Proposition \ref{prop_dist_transition} and Proposition \ref{prop_var_transition}}\label{app_proof_31}

@@

Based on Lemma \ref{lem_hat_v}, we can prove Proposition \ref{prop_dist_transition}.
\begin{proof}[Proof of Proposition \ref{prop_dist_transition}]
	Since $\Sigma_1 = \Sigma_2$, we know that $M = \Sigma_1^{1/2}\Sigma_2^{-1/2} = \hat{v} \id$.
	Using Lemma \ref{lem_minv} and \ref{lem_cov_shift}, we can track the reduction of variance from $\hat{\beta}_t^{\MTL}$ to $\hat{\beta}_t^{\STL}$ as whp
	\begin{align}
		\Delta_{\vari} &\define \sigma^2 \bigbrace{\frac{1}{c_2 - 1} - \frac{1}{c_1 + c_2}\cdot\frac{1}{a_1\hat{v}^2 + a_2} + \bigo{p^{-1/2+\varepsilon}}} \nonumber\\
		&= \sigma^2 \bigbrace{\frac{c_1}{(c_2-1)(c_1 + c_2 -1)} + \bigo{p^{-1/2 + \varepsilon}}},\label{Deltavar}
	\end{align}
	where we use equation \eqref{eq_a12} and Lemma \ref{lem_hat_v}.
	Next we consider the model shift bias
	\begin{align*}
		\Delta_{\beta} &\define \hat{v}^2 \bignorm{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_s - \hat{v}\beta_t)}^2 \\
		&= d^2 \cdot \bignormFro{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1}^2 + \bigo{p^{-1/2 + \varepsilon}d^2}
	\end{align*}
	Using Lemma \ref{prop_model_shift_tight}, we get an upper and lower bound on $\Delta_{\beta}$ as
	\be\label{Deltabeta} \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 \le \Delta_{\beta} / \bigbrace{p\cdot d^2 \cdot \frac{c_1^2}{(c_1+c_2)^2} \cdot \frac{1 + a_3 + a_4}{(a_1 + a_2)^2} + \bigo{p^{1/2+\e}d^2} } \le \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4. \ee
	By solving equations \eqref{eq_a12}, \eqref{eq_a3} and \eqref{eq_a4}, we get
	\begin{align*}
		& a_1 = \frac{c_1(c_1 + c_2 - 1)}{(c_1 + c_2)^2} + O(p^{-1/2 + \varepsilon}),
		a_2 = \frac{c_2(c_1 + c_2 - 1)}{(c_1 + c_2)^2} + O(p^{-1/2 + \varepsilon}), \\
		& a_3 = \frac{c_2}{(c_1 + c_2)(c_1 + c_2 - 1)} + O(p^{-1/2 + \varepsilon}),
		a_4 = \frac{c_1}{(c_1 + c_2)(c_1 + c_2 - 1)} + O(p^{-1/2 + \varepsilon}).
	\end{align*}
	Hence we obtain that
	\begin{align*}
		\frac{1 + a_3 + a_4}{(a_1 + a_2)^2}
		= \frac{(c_1 + c_2)^3}{(c_1 + c_2 - 1)^3} + \bigo{p^{-1/2+\varepsilon}}.
	\end{align*}
	To sum up, we have shown that when
	\[ \frac{p\cdot d^2}{\sigma^2} < \frac{(c_1 + c_2 - 1)^2}{c_1(c_1 + c_2) (c_2 - 1)} \cdot \bigbrace{1 + \sqrt{\frac 1 {c_1}}}^{-4}, \]
	we have that $\Delta_{\vari} > \Delta_{\beta}$, which implies that $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$.
	When
	\[ \frac{p\cdot d^2}{\sigma^2} > \frac{(c_1 + c_2 - 1)^2}{c_1 (c_1 + c_2) (c_2 - 1)} \cdot \bigbrace{1 - \sqrt{\frac 1 {c_1}}}^{-4}, \]
	we have that $\Delta_{\vari} < \Delta_{\beta}$, which implies that $\te(\hat{\beta}_t^{\MTL})$ is higher than $\te(\hat{\beta}_t^{\STL})$.
	The result follows by taking $c_1 = c_2 = c$ for the above equation.
\end{proof}

@@

Recall that under the setting of Example \ref{ex_basic}, \eqref{Deltavar} and \eqref{Deltabeta} hold.  Then we get that $\te(\hat{\beta}_t^{\MTL})< \te(\hat{\beta}_t^{\STL})$ whp if
\be\label{pos1}\left(pd^2 \cdot \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} + \bigo{p^{1/2+\e}d^2} \right)\cdot \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 <  \sigma^2 \bigbrace{1 + \bigo{p^{-1/2 + \varepsilon}}};\ee
otherwise if
\be\label{neg1}\left(pd^2 \cdot \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} + \bigo{p^{1/2+\e}d^2} \right)\cdot \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 >  \sigma^2 \bigbrace{1 + \bigo{p^{-1/2 + \varepsilon}}},\ee
then we have $\te(\hat{\beta}_t^{\MTL})> \te(\hat{\beta}_t^{\STL})$ whp.

Now we prove the first statement of Proposition \ref{prop_data_size}. Notice that the function
$$ \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} =(c_2-1) \left(1 +\frac{c_2-2}{c_1}+\frac{1}{c_1(c_1+c_2)}\right)^{-1} $$
is strictly increasing with respect to $c_1$ as long as $c_2\ge 3$. In particular, for $c_1\ge 100$, we get the bound by taking $c_1\to \infty$:
$$\frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} \cdot \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 < 1.1^4 (c_2-1).$$
 Hence we conclude that for $c_1\ge 100$, we always have $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ whp as long as $pd^2 \le \frac{2\sigma^2}{3(c_2-1)}$. This gives the first statement.

The second statement can be proved in a similar way. Suppose $c_1 \ge 110$ and $pd^2 > \frac{3\sigma^2}{2(c_2-1)}$. If $c_1 > \frac{(c_2-2) \sigma^2}{2(c_2 - 1) pd^2/3 - \sigma^2}$, we have
\begin{align*}
 pd^2 \cdot \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2}  \cdot \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 > \frac{2}{3} pd^2(c_2-1)\cdot \frac{c_1 }{c_1+(c_2-2)}  >\sigma^2, 
 \end{align*}
 where in the first step we used that 
 $$ \frac{(c_1+c_2)(c_1+c_2-2)}{(c_1+c_2-1)^2}  \cdot \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 > \frac23$$
 for $c_1 \ge 110$. This shows that \eqref{neg1} holds as long as $p$ is large enough, and hence $\te(\hat{\beta}_t^{\MTL})> \te(\hat{\beta}_t^{\STL})$ holds. On the other hand, if $100\le c_1 \le \frac{(c_2-2) \sigma^2}{3(c_2 - 1) pd^2/2 - \sigma^2}$, then we have 
 \begin{align*}
 pd^2 \cdot \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2}  \cdot \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 < \frac{3}{2} pd^2(c_2-1)\cdot \frac{c_1 }{c_1+(c_2-2)}  < \sigma^2, 
 \end{align*}
 where in the first step we used that 
 $$ \frac{(c_1+c_2)(c_1+c_2-2)}{(c_1+c_2-1)^2}  \cdot \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 < \frac32$$
 for $c_1 \ge 100$. This shows that \eqref{pos1} holds as long as $p$ is large enough, and hence $\te(\hat{\beta}_t^{\MTL})< \te(\hat{\beta}_t^{\STL})$ holds.

### 2020-05-16 writeup/transfer.tex

\begin{proposition}\label{prop_data_size}
	In the setting of Example \ref{ex_basic}, assume that $c_1 > ??$ and $c_2  > ??$ are fixed constants.
	We have the following conditions to determine whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$:
	\begin{itemize}
		\item If $d^2 < \frac{\sigma^2}{(c_2 - 1) p}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$, for any $c_1$ larger than $??$.
		\item If $d^2 > \frac{\sigma^2}{(c_2 - 1) p}$, then we have the following transition depending on $c_1$:
		\begin{itemize}
			\item If $?? < c_1 < \frac{c_2 \sigma^2}{(c_2 - 1) pd^2 - \sigma^2}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
			\item If $c_1 > \frac{c_2 \sigma^2}{(c_2 - 1)p d^2 - \sigma^2}$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
		\end{itemize}
	\end{itemize}
	Furthermore, as a function of $c_1$ over the range $[??, \infty]$, the maximum of $\te(\hat{\beta}_t^{\STL}) - \te(\hat{\beta}_t^{\MTL})$ is attained when $c_1 = {c_2\sigma^2}/{\max(2(c_2 - 1)pd^2 -\sigma^2, 0)}$.
\end{proposition}

>>>>

\begin{proposition}\label{prop_data_size}
	In the setting of Example \ref{ex_basic}, assume that $c_1 \ge 100$ and $c_2  \ge 3$ are fixed constants.
	We have the following conditions to determine whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$:
	\begin{itemize}
		\item If $d^2 < \frac{2\sigma^2}{3(c_2 - 1) p}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$, for any $c_1$ larger than $??$.
		\item If $d^2 > \frac{3\sigma^2}{2(c_2 - 1) p}$, then we have the following transition depending on $c_1$:
		\begin{itemize}
			\item If $100 \le c_1 < \frac{(c_2-2)\sigma^2}{3(c_2 - 1) pd^2/2 - \sigma^2}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
			\item If $c_1 > \frac{c_2 \sigma^2}{2(c_2 - 1)p d^2/3 - \sigma^2}$ and $c_1\ge 110$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
		\end{itemize}
	\end{itemize}
	Furthermore, as a function of $c_1$ over the range $[??, \infty]$, the maximum of $\te(\hat{\beta}_t^{\STL}) - \te(\hat{\beta}_t^{\MTL})$ is attained when $c_1 = {c_2\sigma^2}/{\max(2(c_2 - 1)pd^2 -\sigma^2, 0)}$. {\cor (cannot get this because we only have some bounds. If we let $c_1\to \infty$, then the curve of $\te(\hat{\beta}_t^{\STL}) - \te(\hat{\beta}_t^{\MTL})$ already becomes flat, and it is meaningless to discuss the minimum of this function at this point?)}
\end{proposition}
